### 2018-03-26 tex/paper.tex
Fix some typos and add notes

As more than one wag has observed, many of the programs that run on blockchains are neither smart nor indeed form contracts in any real sense. Nonetheless, this terminology has stuck, and has become shorthand for ``programmed interactions on the blockchain''. In an earlier paper we presented a survey of the approaches available~\cite{cryptoeprint:2016:1156}; we give an overview here.

@@

A domain-specific language or DSL is a high-level language designed to work in a specific field or domain. The intention is that because the users will know about the field, the constructs of the language can be designed to be meaningful to them, and also that because of its nature the DSL need not include all the features of a general purpose language. Removing this clutter and having the remaining operations directly reflect the application area is intended to make the language more accessible to domain experts who do not necessarily see themselves as programmers.

@@

The simplest blockchain is Bitcoin. Transactions on Bitcoin are made by spending the (as yet) unspent outputs of previous transactions (`unspent transaction outputs' or UTxOs).The chain need not maintain any state, such as the current value owned by any particular participant: such \emph{account} information is implicit and external to the Bitcoin blockchain. On the other hand, the Ethereum model explicitly keeps track of account values, and so this information needs to be kept as state information on chain. While it is possible to translate between UTxO-based and account-based chains \sjtnote{cite Bruno paper} the UTxO model is more fundamental, and requires less support from the implementation of the chain itself.

@@

Contracts can be conceived of as acting in two ways. In a \emph{push} model, contracts are seen to be pro-active, and to make things happen. In the case of blockchain, this would include making payments or transactions happen. Alternatively, in the \emph{pull} model contracts enable certain things to happen, such as transactions, but the transactions need to be effected by an external player. The pull model makes fewer demands on the blockchain implementation than the push model, which requires the contract to have direct access to modify the chain.

@@

In designing Marlowe we have chosen a number of options; this description applies to Marlowe version 1.2 \sjtnote{Is there an obvious way of making that clear in the repo? I really don't understand how git deals with ``releases'' or versions.}

@@

While earlier versions used a ``push'' model of operations, we have revised the language to give a ``pull'' semantics. This means, for example, that the language can \emph{enable} a payment to be made, but cannot make the payment happen. It is up to the interested party (or parties) to make the change happen. It is therefore possible for the payment not to happen, despite the fact that this is against the interest of the receiving party. The same

@@

The language semantics is described in terms of \emph{commitments} of sums of cryptocurrency -- which we choose to abbreviate to ``cash'' -- and so is agnostic of the underlying implementation.  However, the current redemption mechanism in Plutus Core is not strong enough to support Marlowe, and we are actively investigating the best way to extend this mechanism to implement Marlowe in full on Plutus Core.

@@

The whole contract remains as follows:
\begin{haskellcode}
CommitCash (IdentCC 1) 1
           (ConstMoney 100)
           10 200
           (CommitCash (IdentCC 2) 2
                       (ConstMoney 20)
                       20 200
                       (When (PersonChoseSomething (IdentChoice 1) 1)
                             100
                             (Both (RedeemCC (IdentCC 1) Null)
                                   (RedeemCC (IdentCC 2) Null))
                             (Pay (IdentPay 1) 2 1
                                  (ConstMoney 20)
                                  200
                                  (Both (RedeemCC (IdentCC 1) Null)
                                        (RedeemCC (IdentCC 2) Null))))
                       (RedeemCC (IdentCC 1) Null))
           Null
\end{haskellcode}
In the next section we cover the online contract visualisation and interaction tool for Marlowe.

@@

We have given a concrete model and DSL, namely Marlowe version 1.2. In this section we examine how the model can be extended, how contacts can be analysed, reflect on different design decisions that might be made and discuss some aspects of the implementation on Marlowe within a blockchain environment.

@@

Haskell is a pure functional programming language, and uses the \haskellinline{Monad} type class to encapsulate operations that can potentially have a side effect. Many introductions to monads exist \sjtnote{cite some} but for the sake of this discussion it is enough to think of a value of type \haskellinline{m a} where \haskellinline{m} is a monad as a ``computation that results in something of type \haskellinline{a}''. The \haskellinline{Monad} API allows such computations to be assembled, and then once assembled they may be run, in general causing some side-effects as a part of the computation.

@@

Given that we have a language and a semantics for it, we can examine its properties. These might be properties of individual runs of a particular contract, of all runs of a given contract, or indeed properties of all the contracts of the language. In this section we look at two particularly obvious questions, but these are not the only ones that we might ask.

@@

One possible definition is ``those contracts that do not lead to a \haskellinline{FailedPay} action''.
Given the mechanisms for making and using commitments are limited by timeouts, it should be possible to write contracts that have this property, perhaps with some assumptions about the value of observables.
Looking at this in a little more detail, we would expect the following points to hold.
\begin{itemize}
\item For a \haskellinline{Pay} contract a sufficient balance not to produce a \haskellinline{FailedPay} or a \haskellinline{DuplicateRedeem} action is required.. 
\item For \haskellinline{Both} we require a sufficient balance for both contracts combined. Need to think here whether the two conjuncts should be independent, or whether a commitment in one half can be used in the other.
\item For \haskellinline{Choice} we require a sufficient balance for the maximum of the requirements of the individual contracts. 
\item For \haskellinline{When}, need to ``age'' the existing commitments relative to the new starting point of the continuation contract. 
\item Are there any constraints on  \haskellinline{CommitCash} and \haskellinline{RedeemCC} statements? Not directly, but there need to be sufficient funds available to pay if needed, or enough time for the commitments to time out, if not.
\end{itemize}

@@

That is, for each contract do there exist sequences of values of \haskellinline{Commit} and \haskellinline{OS} such that stepping the contract with these inputs leads to \haskellinline{Null}? Some relevant points include:
\begin{itemize}
\item For \haskellinline{When}, need to be able to make \haskellinline{interpretObs st obs os} true. Since there is a timeout to \haskellinline{When} then there's no need to impose this condition, as we'll progress at timeout if not before.
\item As we have put timeouts on to \haskellinline{CommitCash} and \haskellinline{RedeemCC} operations, there are no remaining requirements on these constructs for contract execution to make progress, albeit eventually.
\item We need to check the status of \haskellinline{Observations}: what if an observation is unsatisfiable?
\item There are no constraints on other constructs.
\end{itemize}

@@

Does the \haskellinline{Both} constructor potentially produce invalid/misleading contracts, combining a commitment in one half with its use in the other? The problem is that the ``use'' conjunct could potentially be executed, leading to a failed payment.

@@

Commitments and values of observables need to be recorded somehow, and this raises the potential issue about bloating the chain with data. In general it will only be necessary to store a signed hash value of the observable, and this will keep data usage bounded.

@@

Thanks to IOHK staff Manuel Chakravarty, Duncan Coutts, Bernardo David, Charles Hoskinson, Aggelos Kiyakis, Darry McAdams, Bruno Woltzenlogel Paleo, and Phil Wadler for their collaboration, and to colleagues from the Universities of Kent and Leicester for their comments on previous presentations of this work.

>>>>

As more than one wag has observed, many of the programs that run on blockchains are neither smart nor indeed form 
contracts in any real sense. Nonetheless, this terminology has stuck, and has become shorthand for ``programmed 
interactions on the blockchain''. In an earlier paper, we presented a survey of the approaches 
available~\cite{cryptoeprint:2016:1156}; we give an overview here.

@@

A domain-specific language or DSL is a high-level language designed to work in a specific field or domain. The intention 
is that because the users will know about the field, the constructs of the language can be designed to be meaningful to 
them, and also that, because of its nature, the DSL need not include all the features of a general purpose language. 
Removing this clutter and having the remaining operations directly reflect the application area is intended to make the 
language more accessible to domain experts who do not necessarily see themselves as programmers.

@@

The simplest blockchain is Bitcoin. Transactions on Bitcoin are made by spending the (as yet) unspent outputs of 
previous transactions (`unspent transaction outputs' or UTxOs). The chain need not maintain any state, such as the 
current value owned by any particular participant: such \emph{account} information is implicit and external to the 
Bitcoin blockchain. On the other hand, the Ethereum model explicitly keeps track of account values, and so this 
information needs to be kept as state information on chain. While it is possible to translate between UTxO-based and 
account-based chains \sjtnote{cite Bruno paper} the UTxO model is more fundamental, and requires less support from the 
implementation of the chain itself.

@@

Contracts can be conceived of as acting in two ways. In a \emph{push} model, contracts are seen to be pro-active, and to 
make things happen. In the case of blockchain, this would include making payments or transactions happen. 
Alternatively, in the \emph{pull} model, contracts enable certain things to happen, such as transactions, but the 
transactions need to be effected by an external player. The pull model makes fewer demands on the blockchain 
implementation than the push model, which requires the contract to be triggered automatically.

@@

In designing, Marlowe we have chosen a number of options; this description applies to Marlowe version 1.2 \sjtnote{Is 
there an obvious way of making that clear in the repo? I really don't understand how git deals with ``releases'' or 
versions.}\plsnote{We can create a branch and point to that.}

@@

While earlier versions used a ``push'' model of operations, we have revised the language to give a ``pull'' semantics. 
This means, for example, that the language can \emph{enable} a payment to be made, but cannot make the payment happen. 
It is up to the interested party (or parties) to make the change happen. It is therefore possible for the payment not to 
happen, despite the fact that this is against the interest of the receiving party.\plsnote{There was a trailing ``the 
same'' here. Did you think of another example?}

@@

The language semantics is described in terms of \emph{commitments} of sums of cryptocurrency -- which we choose to 
abbreviate to ``cash'' or ``money'' -- and so is agnostic of the underlying implementation.  However, the current 
redemption mechanism in Plutus Core is not strong enough to support Marlowe, and we are actively investigating the best 
way to extend this mechanism to implement Marlowe in full on Plutus Core.

@@

The whole contract remains as follows:
\begin{haskellcode}
CommitCash (IdentCC 1) 1
           (ConstMoney 100)
           10 200
           (CommitCash (IdentCC 2) 2
                       (ConstMoney 20)
                       20 200
                       (When (PersonChoseSomething (IdentChoice 1) 1)
                             100
                             (Both (RedeemCC (IdentCC 1) Null)
                                   (RedeemCC (IdentCC 2) Null))
                             (Pay (IdentPay 1) 2 1
                                  (ConstMoney 20)
                                  200
                                  (Both (RedeemCC (IdentCC 1) Null)
                                        (RedeemCC (IdentCC 2) Null))))
                       (RedeemCC (IdentCC 1) Null))
           Null
\end{haskellcode}
In the next section, we cover the online contract visualisation and interaction tool for Marlowe.

@@

We have given a concrete model and DSL, namely Marlowe version 1.2. In this section we examine how the model can be 
extended, how contacts can be analysed, reflect on different design decisions that might be made and discuss some 
aspects of the implementation of Marlowe within a blockchain environment.

@@

Haskell is a pure functional programming language, and uses the \haskellinline{Monad} type class to encapsulate
operations that can potentially have a side effect. Many introductions to monads exist \sjtnote{cite some} but for the 
sake of this discussion it is enough to think of a value of type \haskellinline{m a} where \haskellinline{m} is a monad 
as a ``computation that results in something of type \haskellinline{a}''. The \haskellinline{Monad} API allows 
suchcomputations to be assembled, and then, once assembled, they may be run, in general causing some side-effects as a 
part of the computation.

@@

Given that we have a language and a semantics for it, we can examine its properties. These might be properties of 
individual runs of a particular contract, of all runs of a given contract, or indeed properties of all the contracts of 
the language. In this section, we look at two particularly obvious questions, but these are not the only ones that we 
might ask.

@@

One possible definition is ``those contracts that do not lead to a \haskellinline{FailedPay} action''.
Given the mechanisms for making and using commitments are limited by timeouts, it should be possible to write contracts that have this property, perhaps with some assumptions about the value of observables.
Looking at this in a little more detail, we would expect the following points to hold.
\begin{itemize}
\item For a \haskellinline{Pay} contract a sufficient balance not to produce a \haskellinline{FailedPay}.
\item For \haskellinline{Both} we require a sufficient balance for both contracts combined. Need to think here whether the two conjuncts should be independent, or whether a commitment in one half can be used in the other.
\item For \haskellinline{Choice} we require a sufficient balance for the maximum of the requirements of the individual contracts. 
\item For \haskellinline{When}, need to ``age'' the existing commitments relative to the new starting point of the 
continuation contract.
\item \haskellinline{RedeemCC} should not happen twice in any execution, so as to avoid producing a
\haskellinline{DuplicateRedeem} action.
\item Are there any constraints on \haskellinline{CommitCash} statements? Not directly, but there need to be sufficient 
funds available to pay if needed, and enough time for the commitments to time out.
\end{itemize}

@@

That is, for each contract do there exist sequences of values of \haskellinline{Commit} and \haskellinline{OS} such that stepping the contract with these inputs leads to \haskellinline{Null}? Some relevant points include:
\begin{itemize}
\item For \haskellinline{When}, we need to be able to make \haskellinline{interpretObs st obs os} true. Since there is 
a timeout to \haskellinline{When} then there is no need to impose this condition, as we will progress at timeout if not 
before.
\item As we have put timeouts on to \haskellinline{CommitCash} and \haskellinline{RedeemCC} operations, there are no remaining requirements on these constructs for contract execution to make progress, albeit eventually.
\item We need to check the status of \haskellinline{Observations}: what if an observation is unsatisfiable?
\item There are no constraints on other constructs.
\end{itemize}

@@

Does the \haskellinline{Both} constructor potentially produce invalid/misleading contracts, combining a commitment in 
one half with its use in the other? The problem is that the ``use'' conjunct could potentially be executed, leading to 
a failed payment. \plsnote{What do you mean with the ``use'' conjunct?}

@@

Commitments and values of observables need to be recorded somehow, and this raises the potential issue about bloating 
the chain with data. In general, it will only be necessary to store a signed hash value of the observable, and this will 
keep data usage bounded.

@@

Thanks to IOHK staff Manuel Chakravarty, Duncan Coutts, Bernardo David, Charles Hoskinson, Aggelos Kiyakis, Darryl 
McAdams, Bruno Woltzenlogel Paleo, and Phil Wadler for their collaboration, and to colleagues from the Universities of 
Kent and Leicester for their comments on previous presentations of this work.
